---
title: "Mixed effect model selection"
author: "Wing-Ho Ko"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages

```{r}
library(tidyverse)
library(lme4)
library(lmerTest)
library(varTestnlme)
library(broom)
library(broom.mixed)
```


## Loading dataset

```{r}
fish <- read_csv("data/fishdata.csv")
glimpse(fish)
```

## Determining the fixed effect component

Consider a step-up strategy. We first construct a suitable fixed-effect model as starting point, which we can automate using `step()`

```{r}
fish_lm0 <- lm(Trophic_Pos ~ Fish_Length, fish)
fish_lm1 <- step(fish_lm0, Trophic_Pos ~ Fish_Length * Fish_Species)
```

## Determining the random effect component

So our starting fixed-effect model is the fully interacting model between fish length and fish species. Now consider adding random effect associated with lake

```{r}
fish_lmm1 <- lmer(
  Trophic_Pos ~ Fish_Length * Fish_Species + (1 | Lake),
  fish, 
  REML=FALSE
)
summary(fish_lmm1)
```

Since the difference between `fish_lm1` and `fish_lmm1` is a variance component and at null hypothesis its value lie on the boundary of parameter space, we use `varCompTest()` to compare the two models

```{r}
varCompTest(fish_lmm1, fish_lm1)
```

```{r}
AIC(fish_lmm1, fish_lm1)
```


So the random factor is improving the model. Next we can include additional random effect, for example:

```{r}
fish_lmm2 <- lmer(
  Trophic_Pos ~ Fish_Length * Fish_Species + (Fish_Length || Lake),
  fish, 
  REML=FALSE
)
summary(fish_lmm2)
```

```{r}
fish_lmm3 <- lmer(
  Trophic_Pos ~ Fish_Length * Fish_Species + (Fish_Species | Lake),
  fish, 
  REML=FALSE
)

summary(fish_lmm3)
```

Both of these models struggle numerically. Nevertheless, let's continue checking with `varCompTest()`:

```{r}
varCompTest(fish_lmm2, fish_lmm1)
```

```{r}
AIC(fish_lmm2, fish_lmm1)
```


```{r}
varCompTest(fish_lmm3, fish_lmm1)
```

```{r}
AIC(fish_lmm3, fish_lmm1)
```

So from both `AIC()` and `varCompTest()` we should not include any of the additional random effect terms.


